Factoring RSA keys with Shor's algorithm

### 2020-05-31 paper/paper.tex
paper: Add Shor improvement + application

\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{physics}
\begin{document}

@@

\subsection{Order-finding subroutine}
We can further reduce the order-finding problem to the well-known problem of finding the eigenvalues of a unitary operator, which can be solved with the quantum phase estimation algorithm. Shor exploits the unitary operator
\begin{equation}
U_a: x \rightarrow ax \mod N
\end{equation}
whose eigenvalues\footnote{The reader can prove this by substitution: the eigenvectors take the form $\frac{1}{\sqrt{r}}\sum_{m=0}^{r-1}\exp(\frac{-2\pi i m n}{r})\ket{x^k \mod N}$.} are a function of $r$:
\begin{equation}
e^{2i\pi k/r}\text{ for some integer }k
\end{equation}

@@

We devise a simple client-server application that simulates an email service. Users can log in with their username and password, and once logged in they can browse their email messages. It uses a simple cryptographic protocol that simulates TLS, where the server sends its public key to the client, which in turn sends an encrypted AES key (a "session key") allowing for secure two-way communication. The goal of this demonstration is to capture the public key and factor it, allowing a third party to decrypt messages intended for the server and snoop on the credentials and the email messages.

\subsection{Shor}

@@

\begin{thebibliography}{00}
\bibitem{shor} P. W. Shor, "Algorithms for quantum computation: discrete logarithms and factoring", SIAM J.Sci.Statist.Comput. 26, 1997
\bibitem{rsa} R. Rivest, A. Shamir, L. Adleman, "A Method for Obtaining Digital Signatures and Public-Key Cryptosystems", Communications of the ACM 21, 1978
\bibitem{miller} G. L. Miller, "Riemann's hypothesis and tests for primality", J. Comput. System Sci., 1976
\bibitem{vbe} V. Vedral, A. Barenco, A. Ekert, "Quantum Networks for Elementary Arithmetic Operations", Physical Review A 54
\bibitem{ibm} IBM, "Shor's algorithm",\\https://quantum-computing.ibm.com/docs/guide/q-algos/shor-s-algorithm
\end{thebibliography}
\vspace{12pt}

>>>>

\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{physics}
\begin{document}

@@

\subsection{Order-finding subroutine}\label{order-finding}
We can further reduce the order-finding problem to the well-known problem of finding the eigenvalues of a unitary operator, which can be solved with the quantum phase estimation algorithm. Shor exploits the unitary operator
\begin{equation}
U_a: x \rightarrow ax \mod N
\end{equation}
whose eigenvalues\footnote{The reader can prove this by substitution: the eigenvectors take the form $\frac{1}{\sqrt{r}}\sum_{m=0}^{r-1}\exp(\frac{-2\pi i m n}{r})\ket{x^k \mod N}$.} are a function of $r$:
\begin{equation}
e^{2i\pi k/r}\text{ for some integer }k
\end{equation}

@@

\subsection{Shor}
We use the implementation of Shor's algorithm from the Qiskit library. The implementation uses the techniques outlined in section \ref{order-finding}, where the modular exponentiation circuit used to construct $U_b$ is derived in \cite{vbe}.

We present a dramatic improvement upon this algorithm that brings the time spent constructing the circuit from 127 seconds (Qiskit's implementation at the time of writing) to 6 for $N=33$. The resulting circuit is identical and thus has the same execution time; this improvement merely makes the circuit construction faster.

Simple line profiling shows that the Qiskit implementation spends a large amount of time "composing" subcircuits together (for instance, appending an IQFT circuit as part of the QPE routine). Delving into the library code shows that the composition of circuits is implemented by transforming the operands into DAGs\footnote{DAG: Directed Acyclic Graph.}, merging those, and then retransforming back into a circuit. We report an abridged version:

\lstset{basicstyle=\ttfamily\footnotesize,breaklines=true}
\begin{lstlisting}[language=Python]
def compose(self, other):
        dag_self = circuit_to_dag(self)
        dag_other = circuit_to_dag(other)
        dag_self.compose(dag_other)
        return dag_to_circuit(dag_self)
\end{lstlisting}

Further profiling shows that the conversion from and to DAGs is very computationally intensive with respect to the other steps. The key intuition is that multiple compositions are therefore wasteful, as they spend a lot of time converting intermediate values back and forth; representing them as DAGs instead, only converting to and from DAGs respectively at the beginning and at the end of the algorithm, allows for much more efficient composition. In reality the composition takes additional arguments (most importantly the qubit indices on which to compose), which need to be stored along with the DAG - we call this object a "tagged DAG". This simple modification then allows to construct the circuit correctly with dramatic speed improvements.

This improvement was discovered independently by Thomas Metcalfe, who submitted it in a slightly different form on May 25 as part of pull request \#975 to Qiskit Aqua. As the PR is not yet merged at the time of writing, we have released our implementation \cite{github-repo} as a separate repository.

\subsection{Application}
We develop a simple client-server application that simulates an email service. Users can log in with their username and password, and once logged in they can browse their email messages. It uses a simple cryptographic protocol that simulates TLS, where the server sends its public key to the client, which in turn sends an encrypted AES key (a "session key") allowing for secure two-way communication. The goal of this demonstration is to capture the public key and factor it, allowing a third party to decrypt messages intended for the server and snoop on the credentials and the email messages.

@@

\begin{thebibliography}{00}
\bibitem{shor} P. W. Shor, "Algorithms for quantum computation: discrete logarithms and factoring", SIAM J.Sci.Statist.Comput. 26, 1997
\bibitem{rsa} R. Rivest, A. Shamir, L. Adleman, "A Method for Obtaining Digital Signatures and Public-Key Cryptosystems", Communications of the ACM 21, 1978
\bibitem{miller} G. L. Miller, "Riemann's hypothesis and tests for primality", J. Comput. System Sci., 1976
\bibitem{vbe} V. Vedral, A. Barenco, A. Ekert, "Quantum Networks for Elementary Arithmetic Operations", Physical Review A 54
\bibitem{ibm} IBM, "Shor's algorithm",\\https://quantum-computing.ibm.com/docs/guide/q-algos/shor-s-algorithm
\bibitem{github-repo} G. Muscarello, https://github.com/CapacitorSet/qiskit-fast-shor
\end{thebibliography}
\vspace{12pt}
